---
title: "Warmup02-Jose-Lucar.Rmd"
author: "Jose Lucar"
date: "9/13/2018"
output: html_document
---
```{r}

github <-'https://github.com/ucb-stat133/stat133-fall-2018/' 
repo <-'raw/master/data/tents.RData'
download.file(
url = paste0(github, repo),
destfile = "tents.RData")
load("tents.RData")
```


## classes
```{r}

ls()
class(bestuse)
class(brand)
class(capacity)
class(github)
class(height)
class(price)
class(repo)
class(season)
class(tent)
class(weight)
```

## Is Vector
```{r}
is.vector(bestuse)
is.vector(brand)
is.vector(capacity)
is.vector(github)
is.vector(height)
is.vector(price)
is.vector(repo)
is.vector(season)
is.vector(tent)
is.vector(weight)
```

## Is Factor
```{r}

is.factor(bestuse)
is.factor(brand)
is.factor(capacity)
is.factor(github)
is.factor(height)
is.factor(price)
is.factor(repo)
is.factor(season)
is.factor(tent)
is.factor(weight)

typeof(bestuse)
typeof(brand)
typeof(capacity)
typeof(github)
typeof(height)
typeof(price)
typeof(repo)
typeof(season)
typeof(tent)
typeof(weight)
```
 
## Mode

```{r}
mode(bestuse)
mode(brand)
mode(capacity)
mode(github)
mode(height)
mode(price)
mode(repo)
mode(season)
mode(tent)
mode(weight)
```

## Length

```{r}
length(bestuse)
length(brand)
length(capacity)
length(github)
length(height)
length(price)
length(repo)
length(season)
length(tent)
length(weight)
```

## Sring
```{r}

str(tent)
str(bestuse)
str(brand)
str(capacity)
str(github)
str(height)
str(price)
str(repo)
str(season)
str(tent)
str(weight)
```
## extracting the fisrt fice labels

```{r}
head(tent,n=5)
head(bestuse,n=5)
head(brand,n=5)
head(capacity,n=5)
head(github,n=5)
head(height,n=5)
head(price,n=5)
head(repo,n=5)
head(season,n=5)
head(tent,n=5)
head(weight,n=5)

head(tent,n=3)
```

## Plots, Bars, Xlim, Ylim, Stats...

```{r}


summary(price)

sd(price)
mean(price)
min(price)
max(price)
median(price)
quantile(price)
hist(price)
boxplot(price)
density(price)

plot(density(price))
plot(density(height))
plot(density(weight))

# The plot density grapgh has a bell shaped skewed. 

sd(height)
#the center or the mean is greater than the median because it has a right hand tail.

boxplot(height)
mean(height)
min(height)
max(height)  
sd(height)
median(height)

bestuse
factor(bestuse)
table(bestuse)
table(frequency(bestuse))

table(bestuse)
bestuse_fac<- factor(bestuse)
table(bestuse_fac)
bestuse_fac[1:5]
bestuse_fac
table(bestuse_fac)[1]/length(bestuse_fac)
table(bestuse_fac)[2]/length(bestuse_fac)
table(bestuse_fac)[3]/length(bestuse_fac)

relative_freq_bestuse_fact <-(table(bestuse_fac)[1:3]/length(bestuse_fac)*100)
relative_freq_bestuse_fact

barplot((relative_freq_bestuse_fact)/100)

# 2) Scatter Plot

head(height)
height
head(price)
relation_h_p<-plot(height,price)

x<-height 
y<-price
plot.new()
plot.window(xlim= c(0,250),ylim=c(0,700))
axis(1, col.axis= "blue")
axis(2, col.axis="red" , las= 1)
title( main= "height & price")
points(x, y)
```


```{r}

# A vector hight_vec
height
height_vec <-c(height)


# the lenght of the vector hight_vec
length(height_vec)

# length of height
len_height<-c(length(height))
len_height


# the mean for height
x_mean_height=(sum(height)/len_height)
x_mean_height


#length of price
n_price<-c(length(price))
n_price

#mean of height
x_mean_height=(sum(height)/len_height)
x_mean_height

#mean of price:
y_mean_price=(sum(price)/n_price)
y_mean_price

#Variance of height:

variance=(sum((height-x_mean_height)^2))/ (len_height)
variance

## Using the built-in formula for variance gives an slightly different result than calculating it automatically. 

var(height)

#standard deviation of height: 

(variance)^(1/2)

```


## More Manipulation:

### how many tents have a price less than or equal to $300
```{r}
weight_lbs<-c((weight)* .0022)

height_in<-c((height)*0.393701)

```
### how many tents have a price less than or equal to $300

```{r}
price_equal_less_300 <- length(price[price>=300])

price_equal_less_300


```

### how many tents have a price between $300 and $400 (including both $300 and $400 prices)

```{r}

price_btw_300_400 <- length((price[(price<=400)|(price>=300)]))
price_btw_300_400
```
 
###  what’s the name of the tent with maximum price

```{r}

tent[which.max(price)]

```
### how many tents have a price > $400 AND weight < 1500 grams


```{r}
tent[(price>400) & (weight < 1500)]

```
### calculate the 90th percentile for height and assign it to the object height_p90 (display this value)

``` {r}
height_p90 <- quantile(height, c(.90))
height_p90
quantile(height, probs = .9)

```
### calculate the 90th percentile for weight and assign it to the object weight_p90 (display this value)

```{r}
weight_p90 <- quantile(weight, c(.90))
weight_p90

```
### display the name of the tents with height > height_p90 AND weight > weight_p90

```{r}
last_10_percentile=tent[(height > height_p90) & (weight > weight_p90)]

last_10_percentile

```

intervals  | labels
---------------|--------
     (0, 1000] |  1kg
  (1000, 2000] |  2kg
  (2000, 3000] |  3kg
  (3000, 4000] |  4kg
  (4000, 5000] |  5kg
  (5000, 6000] |  6kg
  (6000, 7000] |  7kg
  (7000, 8000] |  8kg
  (8000, 9000] |  9kg



```{r}
weight

weight_factor <- factor(weight)
weight_factor
str(weight_factor)

season_factor <- factor(season)
season_factor
str(season_factor)
is.ordered(season_factor)

```
Without any doubt 'working with factor' was the most difficult part of all the warmups. I will try to get support from GSI next lab.I even have to delete a lot of my work on this because I wasn't able to knit it.


### Coments and Reflections:

- I think remember how to make vector with different features were hard because there are many ways to create vector with specific type of data.

- I don't think anything was really easy, but it was interesting to have a link to help us make the graphs

- Again, I needed help to complete the assigment and it was basically the same friend that helped with the first warm-up. This person doesn't know almost nothing about r like me, but we had two laptos on and a lot of enthusiasm to find how to solve each question. 

- The most time consuming part was the figuering out how to do the boxplot with ** _title_, _axis_ and _points_ **

- I did not understand too much the link because averything in there is new for me, but the more time I spend reading it again and again, the more I get.

- The super most frustraiting thing was trying to rememver what covariance was. I couldn't do it because I couldn't fully understand it. we basically skipped this part, I believe, in my ** Stat 20 ** class from _last year_




