---
title: "lab05-jose-lucar.Rmd"
author: "Jose Lucar"
date: "9/28/2018"
output: html_document
---
```{r}
library(dplyr)
library(ggplot2)
library(readr)
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, fig.path = '../images/')
```

- _This option to download the data only worked when the data is in your working directory. of course, you can acces the data direcctory from here in order to download a new data_ for example you can use the following when your data is inside the **Data* directory:
    - _dat <- read_csv('../Data/'nba2018-players.csv')_ 

```{r}
dat <- read_csv('../Data/nba2018-players.csv')
```

-  **slice()  allows you to select rows by position**

```{r}
three_rows<- slice(dat, 1:3)
three_rows
```

-  **filter()  allows you to select rows by condition***

```{r}
gt_85<- filter(dat, height > 85)
gt_85
```

-  **select()  allows you to select columns by name***

```{r}
player_height<- select(dat, player,height)
player_height
```

### Your Turn:

-  **use 'slice()' to subset the data by selecting the first 5 rows.**

```{r}
slice(dat, 1:5)
```

-   **use slice() to subset the data by selecting rows 10, 15, 20, …, 50.**

```{r}

sli_rows<- c(seq( from = 10, to = 50, by = 5))
slice(dat, sli_rows)
```

-   **use slice() to subset the data by selecting the last 5 rows.**

```{r}
slice(tail(dat,n=5))
```

-   **use filter() to subset those players with height less than 70 inches tall.**
    
```{r}
filter(dat, height<70)
```

-   **use filter() to subset rows of Golden State Warriors (‘GSW’)**
    
```{r}
filter(dat, team == 'GSW')
```

-   **use filter() to subset rows of GSW centers (‘C’)**
    
```{r}
filter(dat, team == 'GSW' & position == 'C')
```

-   **use filter() and then select(), to subset rows of lakers (‘LAL’), and then display their names.**
    
```{r}
select(filter(dat, team == "LAL"), player)
```

-   **use filter() and then select(), to display the name and salary, of GSW point guards**

```{r}
select(filter(dat, team == "GSW"), player, salary)
```

-   **find how to select the name, age, and team, of players with more than 10 years of experience, making 10 million dollars or less.**

```{r}
select(filter(dat, salary <= 10000000 & experience > 10), player, age, team)
```

-   **find how to select the name, team, height, and weight, of rookie players, 20 years old, displaying only the first five occurrences (i.e. rows)**

```{r}
slice(select(filter(dat, age == "20") , player, team, height, weight)   ,1:5)
```

-   **using the data frame gsw, add a new variable product with the product of height and weight.**

```{r}
gsw <- filter(dat, team == 'GSW')
gsw <- select(gsw, player, height, weight)
gsw <- slice(gsw, c(4, 8, 10, 14, 15))
mutate(gsw, height*weight)

```

-   **create a new data frame gsw3, by adding columns log_height and log_weight with the log transformations of height and weight.**

```{r}
gsw3<-mutate(gsw, log_height=log(height), log_weight=log(weight))
```

-   **use the original data frame to filter() and arrange() those players with height less than 71 inches tall, in increasing order.**

```{r}
arrange(filter(dat, height<71), height)
```

-   **display the name, team, and salary, of the top-5 highest paid players**

```{r}
slice(select(arrange(dat, desc(salary)),  player, team , salary), 1:5)
dat
```

-   **display the name, team, and points3, of the top 10 three-point players**

```{r}
slice(select(arrange(dat, desc(points3)), player, team, points3), 1:10)
dat
```

-   **create a data frame gsw_mpg of GSW players, that contains variables for player name, experience, and min_per_game (minutes per game), sorted by min_per_game (in descending order)**

```{r}
gsw_mpg<- select(arrange(filter(mutate(dat, min_per_game=minutes/games ), team=="GSW"), desc(min_per_game)), player, experience, min_per_game)
```

-   **use summarise() to get the largest height value.**

```{r}
summarise(dat, max(height))
```

-   **use summarise() to get the standard deviation of points3**

```{r}
summarise(dat, std_points3 = sd(points3))
```

-   **use summarise() and group_by() to display the median of three-points, by team.**

```{r}
summarise(group_by(dat, team), three_points = median(points3))
```

-   **display the average triple points by team, in ascending order, of the bottom-5 teams (worst 3pointer teams)**

```{r}
arrange(slice(tail(summarise(group_by(dat, team), aver_3pointers = mean(points3)), n=5)), aver_3pointers)
```

-   **obtain the mean and standard deviation of age, for Power Forwards, with 5 and 10 years (including) years of experience.**

```{r}
summarise(filter(dat, position == "PF" & experience >= 5 & experience<= 10) , mean_age= mean(age) , sd_age= sd(age))
```

### First contact with ggplot()

-   **Use the data frame gsw to make a scatterplot of height and weight.**

```{r}
GSW<- filter(dat, team== "GSW")
ggplot(data= GSW) + geom_point(aes(x=weight, y=height))
```

-   **Find out how to make another scatterplot of height and weight, using geom_text() to display the names of the players.**

```{r}
ggplot(GSW, aes(weight, height, label= player)) +geom_text( check_overlap = TRUE)
```

-   **Get a scatter plot of height and weight, for ALL the warriors, displaying their names with geom_label().**

```{r}
ggplot(GSW, aes(weight, height, label= player)) +geom_label( check_overlap = TRUE)
```

-   **Get a density plot of salary (for all NBA players).**

```{r}
ggplot(dat, aes(x=salary))+geom_density()
```

-  **Get a histogram of points2 with binwidth of 50 (for all NBA players).**

```{r}
ggplot(dat, aes(x=points2))+ geom_histogram(binwidth = 50)
```

-   **Get a barchart of the position frequencies (for all NBA players).**

```{r}
ggplot(dat, aes(x=position)) + geom_bar(stat= 'density')
```

-  **Make a scatterplot of experience and salary of all Centers, and use geom_smooth() to add a regression line.**

```{r}
ggplot(filter(dat, position == 'C'),aes(x = experience, y = salary)) + geom_point() +geom_smooth(method = lm) 
```

-  **Repeat the same scatterplot of experience and salary of all Centers, but now use geom_smooth() to add a loess line (i.e. smooth line).**

```{r}
ggplot(filter(dat, position == 'C'),aes(x = experience, y = salary)) + geom_point() +geom_smooth(method = loess) 

```


-  ** Make scatterplots of `experience` and `salary` faceting by `position`**
```{r}
ggplot(dat, aes(x = experience, y = salary)) +geom_point(alpha = 0.9) +facet_wrap(~ position)
```


-  **Make scatterplots of `experience` and `salary` faceting by `team`**

```{r}
ggplot(dat, aes(x = experience, y = salary)) +geom_point(alpha = 0.7) +facet_wrap(~ team)
```

-  **Make density plots of `age` faceting by `team`**

```{r}
ggplot(dat, aes(x = age)) +geom_density(fill ='gray') +facet_wrap(~ team)
```


-  **Make scatterplots of `height` and `weight` faceting by `position`**

```{r}
ggplot(dat, aes(x = height, y = weight)) +geom_point(alpha = 0.5) +facet_wrap(~ position)
```


- **Make scatterplots of `height` and `weight`, with a 2-dimensional density, geom_density2d(), faceting by `position`**

```{r}
ggplot(dat, aes(x = height, y = weight)) +geom_point(alpha = 0.3) +geom_density2d() +facet_wrap(~ position)

```


-  **Make a scatterplot of `experience` and `salary` for the Warriors, but this time add a layer with `theme_bw()` to get a simpler background**

```{r}

ggplot(filter(dat, team == 'GSW'),aes(x = experience, y = salary)) + geom_point() +theme_bw()
```


-  **Repeat any of the previous plots but now adding a leyer with another theme e.g. `theme_minimal()`, `theme_dark()`, `theme_classic()`**

```{r}
ggplot(filter(dat, team == 'GSW'),aes(x = experience, y = salary)) + geom_point() +theme_dark()
```


